---
title: "Titanic"
author: "Thiago Pires"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Introduction

This article presents some analyzes with information about the passengers of the famous ship [Titanic](https://en.wikipedia.org/wiki/RMS_Titanic) that tragically wrecked in 1912. The data was got in the platform [Kaggle](https://www.kaggle.com/c/titanic). In the figure below are showed the ports (with dates) and the place of shipwreck.

```{r}

require(dplyr)
require(leaflet)

ports <- bind_rows(
  tibble(codeloc = "C", location = "Cherbourg (10-04-1912)", long = -1.6109, lat = 49.6445),
  tibble(codeloc = "Q", location = "Queenstown (11-04-1912)", long = -8.3211, lat = 51.8535),
  tibble(codeloc = "S", location = "Southampton (10-04-1912)", long = -1.4191, lat = 50.7894)
)

ports %>% 
  leaflet() %>% 
  addTiles() %>%
  addCircleMarkers(~long, ~lat, popup = ~location, label = ~location) %>% 
  addMarkers(lng = -49.9408, lat = 41.7258, 
             icon = makeIcon("sinking-ship.jpg", 50, 50),
             label = "Naufrágio (14-04-1912)") %>% 
  addCircleMarkers(lng = -73.9655, lat = 40.6832, popup = "New York", label = "New York", color = "green")

```

## Read data

```{r}

require(magrittr)

train <- titanic::titanic_train

```

## Manipulation

```{r}

# tables and figures numbers

require(captioner)

fig_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")

# character to factor

train %<>% 
  mutate(Survived = factor(x = Survived, labels = c("no", "yes")),
         Pclass = factor(x = Pclass, labels = c("1st", "2nd", "3rd")),
         Sex = factor(x = Sex))

```

## Univariate analysis

### Ticket class (Pclass)

There were more survivors in first class than in the second and third class (`r fig_nums("pclass", display = "cite")`).

```{r}

require(ggplot2)

train %>% count(Pclass, Survived) %>% 
  group_by(Pclass) %>% 
  mutate(total = sum(n)) %>% 
  ggplot(aes(Pclass, n * 100/total, fill = Survived)) + geom_col() +
  labs(x = "Ticket class", y = "%", fill = "Survived", 
       caption = fig_nums("pclass", "Percentual distribution of survivors according to ticket class")) +
  theme_light()


```

The model shows the odds to survive in the third class is `r round((1 - exp(-1.67039))*100, 2)` (`1 - exp(-1.67039)`) percent lower than in the first class (`r table_nums("pclass", display = "cite")`).

```{r}

require(DT)

model <-
  glm(Survived ~ Pclass, family = binomial, data = train)

model %>%
  summary() %$% coefficients %>% 
  datatable(caption = table_nums("pclass", "Model estimation"), options = list(dom = "t")) %>% 
  formatRound(columns = 1:4 , digits = 3)
  
```

### Sex

There were more female survivors than male (`r fig_nums("sex", display = "cite")`).

```{r}

train %>% count(Sex, Survived) %>% 
  group_by(Sex) %>% 
  mutate(total = sum(n)) %>% 
  ggplot(aes(Sex, n * 100/total, fill = Survived)) + geom_col() +
  labs(x = "Sex", y = "%", fill = "Survived", 
       caption = fig_nums("sex", "Percentual distribution of survivors according to sex")) +
  theme_light()

```

The model shows the odds to survive for male is `r round((1 - exp(-2.513710))*100, 2)` (`1 - exp(-2.513710)`) percent lower than for female (`r table_nums("sex", display = "cite")`).

```{r}

model <-
  glm(Survived ~ Sex, family = binomial, data = train)

model %>%
  summary() %$% coefficients %>% 
  datatable(caption = table_nums("sex", "Estimativas do modelo"), options = list(dom = "t")) %>% 
  formatRound(columns = 1:4 , digits = 3)
  
```

## Interaction between Sex and Ticket Class

The interaction between sex and ticket class was significant (`r table_nums("pclass_sex", display = "cite")`), showing that the differences between the probability of survival between men and women is greater when the ticket class improves (`r fig_nums("pclass_sex", display = "cite")`).

```{r}

model <-
  glm(Survived ~ Pclass*Sex, family = binomial, data = train)

model %>%
  summary() %$% coefficients %>% 
  datatable(caption = table_nums("pclass_sex", "Model estimation"), options = list(dom = "t")) %>% 
  formatRound(columns = 1:4 , digits = 3)

```

```{r}

# table with all possibilities

newdata <- 
  expand.grid(Pclass = c("1st", "2nd", "3rd"),
              Sex = c("male", "female")) %>% as_tibble()

# predictions (probability)

newdata %<>% 
  mutate(Pihat = model %>% predict(newdata = newdata, type = "response"))

newdata %>% 
  ggplot(aes(Sex, Pihat, group = Pclass, colour = Pclass)) + geom_line() + geom_point() +
  labs(x = "Sex", y = expression(pi(Survived == yes)), colour = "Ticket Class",
       caption = fig_nums("pclass_sex", "Interaction between sex and ticket class")) +
  theme_light()

```

### Age

There is no difference between the distribution of age according to survived status (`r fig_nums("age", display = "cite")`).

```{r}

train %>% 
  ggplot(aes(Survived, Age)) + geom_boxplot() +
  labs(caption = fig_nums("age", "Distribution of age according to survived status")) +
  theme_light()

```

Fitting a simple model is showed a little age effect on survival probability (`r table_nums("age", display = "cite")`). The increase of one year in age decreases the odds of survival by `r round((1 - exp(-0.0109635))*100, 2)` (`1 - exp(-0.0109635)`) percent.

```{r}

model <-
  glm(Survived ~ Age, family = binomial, data = train)

model %>%
  summary() %$% coefficients %>% 
  datatable(caption = table_nums("age", "Model estimation"), options = list(dom = "t")) %>% 
  formatRound(columns = 1:4 , digits = 3)
  
```

## Interaction between Age and Ticket Class

Below, note that there is no interaction between age and ticket class (p-values > .05)

```{r}

model <-
  glm(Survived ~ Pclass*Age, family = binomial, data = train)

model %>%
  summary() %$% coefficients %>% 
  datatable(caption = table_nums("pclass_age", "Model estimation"), options = list(dom = "t")) %>% 
  formatRound(columns = 1:4 , digits = 3)

```
